---
title: "Breast milk Attachment - Demographics"
author: "LJS"
date: "2024-06-07"
output: html_document
---

# Libraries 

```{r setup, include=FALSE}

library("rlang") #toolbox
library("dplyr") #data manipulation
library("here") #toolbox
library("tidyverse") #data manipulation
library("tidytext") #data manipulation
library("knitr") #rmarkdown
library("htmltools") #rmarkdown
library("markdown") #rmarkdown
library("httpuv") #rmarkdown
library("vtable") #summary tables
library("gtsummary") #summary tables
library("NCmisc") #toolbox
library("stringr") #data manipulation
knitr::opts_chunk$set(warning = FALSE, message = FALSE) #do not print warning messages

```

# Breast milk exposure during neonatal care and infant attachment

Avoid scientific notation.

```{r}

options(scipen = 999)

```

## Section I: Prepare data

### Merge individual spreadsheets

```{r}

BM_variables <- read.csv(here("raw_data", "IDs_BM.csv"), header=T, dec=".",sep=",")
Attachment_variables <- read.csv(here("raw_data", "IDs_WT_rawtimes_filtered.csv"), header=T, dec=".",sep=",")
Covariate_variables <- read.csv(here("raw_data", "IDs_covariates.csv"), header=T, dec=".",sep=",")
TEA_feeding <- read.csv(here("raw_data", "IDs_TEA_feeding.csv"), header=T, dec=".",sep=",")

Data <- merge(BM_variables, Attachment_variables, by="ID")
Data <- merge(Covariate_variables, Data, by="ID")
Data <- merge(TEA_feeding, Data, by="ID")
rownames(Data) <- Data$ID
Data$ID <- NULL

```

### Breast milk

Calculate the % of exclusively mother breast milk, % of exclusively donor breast milk, % of both mother+donor breast milk and % of exclusively any (mother or donor) breast milk during neonatal care.

```{r}

Data <- Data %>% mutate(prop_days_excl_mother_breastmilk = (days_excl_mbm)/days_nnu)
Data <- Data %>% mutate(prop_days_excl_donor_breastmilk = (days_excl_debm)/days_nnu)
Data <- Data %>% mutate(prop_days_excl_motherdonor_breastmilk = (days_excl_mbm_debm)/days_nnu)
Data <- Data %>% mutate(prop_days_excl_any_breastmilk = (days_excl_mbm + days_excl_debm + days_excl_mbm_debm)/days_nnu)
Data <- Data %>% mutate(prop_days_formula = (days_excl_formula)/days_nnu)
Data <- Data %>% mutate(breastmilk_75_term = case_when((Prematurity==0)~2,
(Prematurity==1 & prop_days_excl_any_breastmilk < 0.75)~0,
(Prematurity==1 & prop_days_excl_any_breastmilk >= 0.75)~1))
Data=Data[complete.cases(Data$breastmilk_75_term),]

```

### Attachment

Categorise infants as secure or insecure based on attachment style. 
Remove participants with an unscorable attachment style (n = 7).

```{r}

Data$AttachmentSecurity = "Secure"

for(i in 1:nrow(Data)){
  if(Data$AttachmentStyle[i]=="Avoidant"){
    Data$AttachmentSecurity[i] <- "Insecure"}
  if(Data$AttachmentStyle[i]=="Resistant"){
    Data$AttachmentSecurity[i] <- "Insecure"}
  if(Data$AttachmentStyle[i]=="Disorganised"){
    Data$AttachmentSecurity[i] <- "Insecure"}
  if(Data$AttachmentStyle[i]=="NotFit"){
    Data$AttachmentSecurity[i] <- "NotFit"}}

Data = Data[Data$AttachmentSecurity=="Secure" | Data$AttachmentSecurity=="Insecure",]

```

### Transform variables to factors

```{r}

factor_columns <- c("breastmilk_75_term", "Sex", "Prematurity", "Exclusive_BF_4.5",
                    "Postnatal_depression_4.5", "Early_onset_sepsis", 
                    "Late_onset_sepsis", "Medical_or_surgical_nec", 
                    "bronchopulmonary_dysplasia", "retinopathy_or_prematurity")

for (col in factor_columns) {
  Data[[col]] <- as.factor(Data[[col]])
}

```

## Section II: Demographics of study sample

### Compare % of exclusively mother, donor or combined breastmilk intake during neonatal care across breastmilk exposure groups

```{r}

Data_milktype <- Data[,grep("breastmilk_75_term|prop_days_excl_mother|prop_days_excl_donor", colnames(Data))]
Data_milktype <- Data_milktype %>% filter(breastmilk_75_term %in% c("0", "1"))

tbl_summary(
  Data_milktype,
  by = breastmilk_75_term,
  label = NULL,
  statistic = list(c("prop_days_excl_mother_breastmilk", "prop_days_excl_donor_breastmilk", "prop_days_excl_motherdonor_breastmilk") ~ c("{median} ({min}, {max})")),
  digits = all_continuous()~3,
  type = NULL,
  value = NULL,
  missing = NULL,
  missing_text = NULL,
  sort = NULL,
  percent = NULL,
  include = everything()
)

```

# Breastmilk data distribution

```{r}

g <- ggplot(Data, aes(x = prop_days_excl_any_breastmilk)) +
  geom_histogram(binwidth = 0.1, fill = "steelblue", color = "black", alpha = 0.7) +
  labs(x = "Proportion of days in the NICU of exclusive\nbreast milk (mother- or donor- expressed)",
       y = "Number of participants") +
  theme_classic() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text = element_text(size = 12)
  )

g

ggsave(here("results/figures", "Prop.NICUdays.exclusiveBM.pdf"), width = 130, height = 100, units = "mm", plot = g, dpi=300)

```

# Feeding type at neonatal visit

```{r}

Data_feeding <- Data[,grep("breastmilk_75_term|visit_", colnames(Data))]

tbl_summary(
  Data_feeding,
  by = breastmilk_75_term,
  label = NULL,
 statistic = list(c("visit_CGA") ~ c("{mean} ({min}, {max})"), 
                  all_categorical() ~ "{n} ({p}%)"),
  digits = all_continuous()~3,
  type = NULL,
  value = NULL,
  missing = NULL,
  missing_text = NULL,
  sort = NULL,
  percent = NULL,
  include = everything()
)

```

# Caregivers

```{r}

Data_caregivers <- Data[,grep("breastmilk_75_term|Caregiver", colnames(Data))]

tbl_summary(
  Data_caregivers,
  by = breastmilk_75_term,
  label = NULL,
  digits = all_continuous()~3,
  type = NULL,
  value = NULL,
  missing = NULL,
  missing_text = NULL,
  sort = NULL,
  percent = NULL,
  include = everything()
)

```

# Descriptive statistics

```{r}

# data wrangling 

Data_baseline <- Data[,grep("Sex|GA|BW|breastmilk_75_term|SIMD|days_nnu|Eth|prop_days_excl_any_breastmilk|prop_days_formula|Early_onset_sepsis|Late_onset_sepsis|Medical_or_surgical_nec|bronchopulmonary_dysplasia|retinopathy_or_prematurity|Postnatal_depression_4.5|Exclusive_BF_4.5", colnames(Data))]

Data_baseline <- Data_baseline[,-grep("visit_CGA", colnames(Data_baseline))]

# create table 

table_dem <- tbl_summary(
  Data_baseline,
  by = breastmilk_75_term,
  label = NULL,
 statistic = list(c("GA", "BW", "BW_zscore", "SIMD", "days_nnu") ~ c("{median} ({min}, {max})"), 
                  all_categorical() ~ "{n} ({p}%)"),
  digits = list(all_categorical()~2, all_continuous()~3),
  type = NULL,
  value = NULL,
  missing = NULL,
  missing_text = NULL,
  sort = NULL,
  percent = NULL,
  include = everything()
)

table_dem 

# export table 

table_dem <- as_tibble(table_dem)
colnames(table_dem) <- c("Characteristic", "Preterm with low BM (N = 28)", "Preterm with high BM (N = 49)", "Term (N = 71)")

write.csv(table_dem, here("results/demographics", "Dem.table.csv"), row.names=FALSE)

# comparisons between preterm infants with low vs high breast milk exposure during neonatal care

## data wrangling 

Data_baseline_pret = Data_baseline[Data_baseline$breastmilk_75_term=="0" | Data_baseline$breastmilk_75_term=="1",]
Data_baseline_pret$breastmilk_75_term = factor(Data_baseline_pret$breastmilk_75_term)

Pval_fun <- function(x) style_pvalue(x, digits = 3)
Stat_fun <- function(x) style_sigfig(x, digits=3)

## create table with comparisons

table_dem_pret <- Data_baseline_pret %>%
  tbl_summary(
    by = breastmilk_75_term,
    statistic = list(
      c("GA", "BW", "BW_zscore", "SIMD", "days_nnu", "prop_days_excl_any_breastmilk", "prop_days_formula") ~ "{median} ({min}, {max})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    digits = list(all_categorical()~2, all_continuous()~3)
  ) %>%
  add_p(
    test = list(
      c("GA", "BW", "BW_zscore", "SIMD", "days_nnu", "prop_days_excl_any_breastmilk", "prop_days_formula") ~ "wilcox.test",
      all_categorical() ~ "chisq.test"
    ),
    pvalue_fun = Pval_fun
  ) %>%
  separate_p_footnotes() %>%
  modify_header(list(statistic ~ "**t-statistic**", p.value ~ "**p-value**")) %>%
  modify_fmt_fun(list(statistic ~ Stat_fun, p.value ~ Pval_fun))

table_dem_pret

## export table

table_dem_pret <- as_tibble(table_dem_pret)
colnames(table_dem_pret) <- c("Characteristic", "Preterm with low BM (N = 28)", "Preterm with high BM (N = 49)", "t-statistic", "p-value")

write.csv(table_dem_pret, here("results/demographics", "Dem.table.preterm.csv"), row.names=FALSE)

```


```{r pressure, echo=FALSE}
markdownToHTML("BM_Attachment_dem.Rmd",output ="BM_Attachment_dem.html")
```

