---
title: "BM_Attachment_sens"
author: "LJS"
date: "2024-07-01"
output: html_document
editor_options: 
  chunk_output_type: inline
---

# Libraries 

```{r}

library("rlang") #toolbox
library("dplyr") #data manipulation
library("here") #toolbox
library("tidyverse") #data manipulation
library("tidytext") #data manipulation
library("knitr") #rmarkdown
library("htmltools") #rmarkdown
library("markdown") #rmarkdown
library("httpuv") #rmarkdown
library("vtable") #summary tables
library("gtsummary") #summary tables
library("stringr") #data manipulation
library("car") #levene tests, qqplot
library("rstatix") #kruskal_effsize, wilcox_effsize
library("nnet") #multinom
library("ggplot2") #plotting
library("performance") #check models performance
library("easystats") #check models performance
library("see") #check models performance

```

# Breast milk exposure during neonatal care and infant attachment

Avoid scientific notation.

```{r}

options(scipen = 999)

```

## Section I: Prepare data

### Merge individual spreadsheets

```{r}

BM_variables <- read.csv(here("raw_data", "IDs_BM.csv"), header=T, dec=".",sep=",")
Attachment_variables <- read.csv(here("raw_data", "IDs_WT_rawtimes_filtered.csv"), header=T, dec=".",sep=",")
Covariate_variables <- read.csv(here("raw_data", "IDs_covariates.csv"), header=T, dec=".",sep=",")
TEA_feeding <- read.csv(here("raw_data", "IDs_TEA_feeding.csv"), header=T, dec=".",sep=",")

Data <- merge(BM_variables, Attachment_variables, by="ID")
Data <- merge(Covariate_variables, Data, by="ID")
Data <- merge(TEA_feeding, Data, by="ID")
rownames(Data) <- Data$ID
Data$ID <- NULL

```

### Breast milk

Calculate the % of exclusively mother breast milk, % of exclusively donor breast milk, % of both mother+donor breast milk and % of exclusively any (mother or donor) breast milk during neonatal care.

```{r}

Data <- Data %>% mutate(prop_days_excl_mother_breastmilk = (days_excl_mbm)/days_nnu)
Data <- Data %>% mutate(prop_days_excl_donor_breastmilk = (days_excl_debm)/days_nnu)
Data <- Data %>% mutate(prop_days_excl_motherdonor_breastmilk = (days_excl_mbm_debm)/days_nnu)
Data <- Data %>% mutate(prop_days_excl_any_breastmilk = (days_excl_mbm + days_excl_debm + days_excl_mbm_debm)/days_nnu)
Data <- Data %>% mutate(prop_days_formula = (days_excl_formula)/days_nnu)
Data <- Data %>% mutate(breastmilk_75_term = case_when((Prematurity==0)~2,
(Prematurity==1 & prop_days_excl_any_breastmilk < 0.75)~0,
(Prematurity==1 & prop_days_excl_any_breastmilk >= 0.75)~1))
Data=Data[complete.cases(Data$breastmilk_75_term),]

```

### Attachment

Categorise infants as secure or insecure based on attachment style. 
Remove participants with an unscorable attachment style (n = 7).

```{r}

Data$AttachmentSecurity = "Secure"

for(i in 1:nrow(Data)){
  if(Data$AttachmentStyle[i]=="Avoidant"){
    Data$AttachmentSecurity[i] <- "Insecure"}
  if(Data$AttachmentStyle[i]=="Resistant"){
    Data$AttachmentSecurity[i] <- "Insecure"}
  if(Data$AttachmentStyle[i]=="Disorganised"){
    Data$AttachmentSecurity[i] <- "Insecure"}
  if(Data$AttachmentStyle[i]=="NotFit"){
    Data$AttachmentSecurity[i] <- "NotFit"}}

Data = Data[Data$AttachmentSecurity=="Secure" | Data$AttachmentSecurity=="Insecure",]

```

### Transform variables to factors

```{r}

factor_columns <- c("breastmilk_75_term", "Sex", "Prematurity", "Exclusive_BF_4.5",
                    "Postnatal_depression_4.5", "Early_onset_sepsis", 
                    "Late_onset_sepsis", "Medical_or_surgical_nec", 
                    "bronchopulmonary_dysplasia", "retinopathy_or_prematurity")

for (col in factor_columns) {
  Data[[col]] <- as.factor(Data[[col]])
}

```

## Section II: Analyses of attachment behaviours controlling for bronchopulmonary dysplasia

```{r}

Data_pret = Data[Data$breastmilk_75_term=="0" | Data$breastmilk_75_term=="1",]
Data_pret$breastmilk_75_term <- as.factor(Data_pret$breastmilk_75_term)

```

### Make a dataframe for results

```{r}

results_table <- data.frame(matrix(ncol = 4, nrow = 4))
x <- c("measure", "effsize", "SE", "p-value")
colnames(results_table) <- x

results_table$measure <- c("HD_Raw", "FNF_Raw", "AA_Raw", "Secure/Insecure")

```

### Breast milk exposure ~ behavioural scales

Run models and create table with results.

```{r}

Data_pret[,grep('HD_Raw|FNF_Raw|AA_Raw', colnames(Data_pret))] <- scale(Data_pret[,grep('HD_Raw|FNF_Raw|AA_Raw', colnames(Data_pret))])

targetdata=Data_pret
colns_oi = colnames(targetdata)[grep('HD|FNF|AA',colnames(targetdata))]

for(i in colns_oi){
  formula.torun = paste0(i, '~bronchopulmonary_dysplasia+breastmilk_75_term')
  fit=lm(as.formula(formula.torun),data=targetdata) 
  report=summary(fit)
  single.model=report$coefficients[nrow(report$coefficients),]
  
  if (i==colns_oi[1]){
    total.report=single.model
  }else{
    total.report=rbind(total.report,single.model)
  }
  
}

rownames(total.report)=colns_oi
Total.report.beh = data.frame(total.report)
colnames(Total.report.beh) <- c("Beta", "SE", "t-value", "p-value")
Total.report.beh$Behaviour <- rownames(Total.report.beh)
rownames(Total.report.beh) <- NULL
new_order = c("Behaviour","Beta", "SE", "t-value", "p-value")
Total.report.beh <- Total.report.beh[, new_order]
Results.report <- Total.report.beh

results_table$`effsize`[which(results_table$measure=="HD_Raw")] <- Results.report$Beta[which(Results.report$Behaviour=="HD_Raw")]
results_table$SE[which(results_table$measure=="HD_Raw")] <- Results.report$SE[which(Results.report$Behaviour=="HD_Raw")]
results_table$"p-value"[which(results_table$measure=="HD_Raw")] <- Results.report$"p-value"[which(Results.report$Behaviour=="HD_Raw")]

results_table$`effsize`[which(results_table$measure=="FNF_Raw")] <- Results.report$Beta[which(Results.report$Behaviour=="FNF_Raw")]
results_table$SE[which(results_table$measure=="FNF_Raw")] <- Results.report$SE[which(Results.report$Behaviour=="FNF_Raw")]
results_table$"p-value"[which(results_table$measure=="FNF_Raw")] <- Results.report$"p-value"[which(Results.report$Behaviour=="FNF_Raw")]

results_table$`effsize`[which(results_table$measure=="AA_Raw")] <- Results.report$Beta[which(Results.report$Behaviour=="AA_Raw")]
results_table$SE[which(results_table$measure=="AA_Raw")] <- Results.report$SE[which(Results.report$Behaviour=="AA_Raw")]
results_table$"p-value"[which(results_table$measure=="AA_Raw")] <- Results.report$"p-value"[which(Results.report$Behaviour=="AA_Raw")]

```

Model diagnostics for Breast milk exposure ~ behavioural scales

```{r}

model_list=list()
for(i in colns_oi){
  formula.torun = paste0(i, '~bronchopulmonary_dysplasia+breastmilk_75_term')
  fit=lm(as.formula(formula.torun),data=targetdata)
  model_list[[i]]=fit}

lapply(model_list, check_model)

#Double-check normality of residuals with Q-Q plots
for (i in model_list) {
  plot(i, which=2)}

```

## Section III: Analyses of secure vs insecure attachment classification controlling for bronchopulmonary dysplasia

```{r}

Data_pret$breastmilk_75_term2 <- relevel(factor(Data_pret$breastmilk_75_term), ref="0")
Data_pret$AttachmentSecurity2 <- relevel(factor(Data_pret$AttachmentSecurity), ref="Insecure")

multinom_test <- multinom(AttachmentSecurity2 ~ breastmilk_75_term2, data = Data_pret)
summary(multinom_test)
z <- summary(multinom_test)$coefficients/summary(multinom_test)$standard.errors
z
p <- (1 - pnorm(abs(z), 0, 1)) * 2
p
exp(coef(multinom_test))

# Main p-values to be corrected for analyses of attachment behaviours and secure vs insecure attachment classification are labelled "pval_1_" .


results_table$`effsize`[which(results_table$measure=="Secure/Insecure")] <- summary(multinom_test)$coefficients[2]
results_table$SE[which(results_table$measure=="Secure/Insecure")] <- summary(multinom_test)$standard.errors[2]
results_table$"p-value"[which(results_table$measure=="Secure/Insecure")] <- p[2]

```

## Section IV: P-value adjustment and save results

```{r}

write.csv(results_table, here("results/comparisons", "Comp.3group.unadj.sens.csv"), row.names = FALSE)

results_table_adjusted = results_table%>%mutate(across(contains("p-value"),~ p.adjust(.x, method = "BH", n = 4)))
write.csv(results_table_adjusted, here("results/comparisons", "Comp.3group.adj.sens.csv"), row.names = FALSE)

```


```{r pressure, echo=FALSE}
markdownToHTML("BM_Attachment_sens.Rmd",output ="BM_Attachment_sens.html")
```

