---
title: "BM_Attachment_cov"
author: "LJS"
date: "2024-07-02"
output: html_document
---

# Libraries 

```{r}

library("rlang") #toolbox
library("dplyr") #data manipulation
library("here") #toolbox
library("tidyverse") #data manipulation
library("tidytext") #data manipulation
library("knitr") #rmarkdown
library("htmltools") #rmarkdown
library("markdown") #rmarkdown
library("httpuv") #rmarkdown
library("vtable") #summary tables
library("gtsummary") #summary tables
library("stringr") #data manipulation

```

# Breast milk exposure during neonatal care and infant attachment

Avoid scientific notation.

```{r}

options(scipen = 999)

```

## Section I: Prepare data

### Merge individual spreadsheets

```{r}

BM_variables <- read.csv(here("raw_data", "IDs_BM.csv"), header=T, dec=".",sep=",")
Attachment_variables <- read.csv(here("raw_data", "IDs_WT_rawtimes_filtered.csv"), header=T, dec=".",sep=",")
Covariate_variables <- read.csv(here("raw_data", "IDs_covariates.csv"), header=T, dec=".",sep=",")
TEA_feeding <- read.csv(here("raw_data", "IDs_TEA_feeding.csv"), header=T, dec=".",sep=",")

Data <- merge(BM_variables, Attachment_variables, by="ID")
Data <- merge(Covariate_variables, Data, by="ID")
Data <- merge(TEA_feeding, Data, by="ID")
rownames(Data) <- Data$ID
Data$ID <- NULL

```

### Breast milk

Calculate the % of exclusively mother breast milk, % of exclusively donor breast milk, % of both mother+donor breast milk and % of exclusively any (mother or donor) breast milk during neonatal care.

```{r}

Data <- Data %>% mutate(prop_days_excl_mother_breastmilk = (days_excl_mbm)/days_nnu)
Data <- Data %>% mutate(prop_days_excl_donor_breastmilk = (days_excl_debm)/days_nnu)
Data <- Data %>% mutate(prop_days_excl_motherdonor_breastmilk = (days_excl_mbm_debm)/days_nnu)
Data <- Data %>% mutate(prop_days_excl_any_breastmilk = (days_excl_mbm + days_excl_debm + days_excl_mbm_debm)/days_nnu)
Data <- Data %>% mutate(prop_days_formula = (days_excl_formula)/days_nnu)
Data <- Data %>% mutate(breastmilk_75_term = case_when((Prematurity==0)~2,
(Prematurity==1 & prop_days_excl_any_breastmilk < 0.75)~0,
(Prematurity==1 & prop_days_excl_any_breastmilk >= 0.75)~1))
Data=Data[complete.cases(Data$breastmilk_75_term),]

```

### Attachment

Categorise infants as secure or insecure based on attachment style. 
Remove participants with an unscorable attachment style (n = 7).

```{r}

Data$AttachmentSecurity = "Secure"

for(i in 1:nrow(Data)){
  if(Data$AttachmentStyle[i]=="Avoidant"){
    Data$AttachmentSecurity[i] <- "Insecure"}
  if(Data$AttachmentStyle[i]=="Resistant"){
    Data$AttachmentSecurity[i] <- "Insecure"}
  if(Data$AttachmentStyle[i]=="Disorganised"){
    Data$AttachmentSecurity[i] <- "Insecure"}
  if(Data$AttachmentStyle[i]=="NotFit"){
    Data$AttachmentSecurity[i] <- "NotFit"}}

Data = Data[Data$AttachmentSecurity=="Secure" | Data$AttachmentSecurity=="Insecure",]

```

### Transform variables to factors

```{r}

factor_columns <- c("breastmilk_75_term", "Sex", "Prematurity", "Exclusive_BF_4.5",
                    "Postnatal_depression_4.5", "Early_onset_sepsis", 
                    "Late_onset_sepsis", "Medical_or_surgical_nec", 
                    "bronchopulmonary_dysplasia", "retinopathy_or_prematurity")

for (col in factor_columns) {
  Data[[col]] <- as.factor(Data[[col]])
}

```

## Section II: Covariate analyses for attachment behaviours

### Categorical variables - bronchopulmonary dysplasia

```{r}

# data wrangling

Data_pret = Data[Data$breastmilk_75_term=="0" | Data$breastmilk_75_term=="1",]
Data_pret$breastmilk_75_term = factor(Data_pret$breastmilk_75_term)

# check data distribution 

targetdata=Data_pret[,grep('HD_Raw|FNF_Raw|AA_Raw', colnames(Data_pret))]
targetdata=as.data.frame(targetdata)
apply(targetdata,2,shapiro.test)

for (cols in colnames(targetdata)) {
  targetdata %>% pull(cols) %>% hist(main = cols)
}

for (cols in colnames(targetdata)) {
  targetdata %>% pull(cols) %>% qqnorm(main = cols)
}

```

In preterm infants, the Happy-Distressed scale is normally distributed, while the Not fretful-Fretful and Attentive-Avoidant scales are non-normally distributed.

Display tables of descriptive statistics and tests by confounder.

```{r}

Pval_fun <- function(x) style_pvalue(x, digits = 3)
Stat_fun <- function(x) style_sigfig(x, digits=3)

```

Associations between bronchopulmonary dysplasia and attachment behaviours.

```{r}

# data wrangling 

targetdata=Data_pret[,grep('broncho|HD_Raw|FNF_Raw|AA_Raw', colnames(Data_pret))]

# create table 

table_cov <- targetdata %>% tbl_summary(by = bronchopulmonary_dysplasia,
  label = NULL,
  statistic = list(c("FNF_Raw", "AA_Raw") ~ c("{median} ({min}, {max})"), c("HD_Raw") ~ c("{mean} ({sd})"), all_categorical() ~ "{n} ({p}%)"),
  digits = everything()~3) %>% 
  add_p(
  test = list(c("FNF_Raw", "AA_Raw") ~ "wilcox.test",
                   c("HD_Raw") ~ "t.test", all_categorical() ~ "chisq.test"),
  pvalue_fun = Pval_fun) %>% 
  separate_p_footnotes() %>% 
  modify_header(
    list(
      statistic ~ "**t-statistic**",
      p.value ~ "**p-value**"
    )
  ) %>%
  modify_fmt_fun(
    list(
      statistic ~ Stat_fun,
      p.value ~ Pval_fun
    ))

# export table

corr_cat_attachbeh <- table_cov
corr_cat_attachbeh <- as_tibble(corr_cat_attachbeh)
colnames(corr_cat_attachbeh) <- c("Characteristic", "Absence of BPD", "Presence of BPD", "t-statistic", "p-value")

write.csv(corr_cat_attachbeh, here("results/covariates", "AttachBeh.cat.csv"), row.names=FALSE)

```

### Continuous variables - BW, GA, SIMD, days in neonatal care (days_nnu)

```{r}

# check data distribution 

targetdata=Data_pret[,grep('BW|GA|SIMD|days_nnu', colnames(Data_pret))]
targetdata=targetdata[,-grep('visit_|BW_zscore', colnames(targetdata))]
targetdata=as.data.frame(targetdata)
apply(targetdata,2,shapiro.test)

for (cols in colnames(targetdata)) {
  targetdata %>% pull(cols) %>% hist(main = cols)
}

for (cols in colnames(targetdata)) {
  targetdata %>% pull(cols) %>% qqnorm(main = cols)
}

```

In preterm infants, BW is normally distributed, while gestational age at birth (GA), socioeconomic status (SIMD), and days in neonatal care (days_nnu) are non-normally distributed.

Associations between demographic/clinical variables and attachment behaviours.

```{r}

# spearman correlations

targetdata=Data_pret[,grep('BW|GA|SIMD|days_nnu|HD_|FNF_|AA_', colnames(Data_pret))]
targetdata=targetdata[,-grep('visit_|BW_zscore', colnames(targetdata))]

cors <- function(df) { 
  # turn all three matrices (r, n, and P into a data frame)
  M <- Hmisc::rcorr(as.matrix(df), type="spearman")
  # return the three data frames in a list return(Mdf)
  Mdf <- map(M, ~data.frame(.x))
}

newdata <- cors(targetdata)

formatted_cors <- function(df){
  cors(df) %>%
    map(~rownames_to_column(.x, var="measure1")) %>%
    map(~pivot_longer(.x, -measure1, names_to="measure2")) %>% 
    bind_rows(.id = "id") %>%
    tidyr::pivot_wider(names_from = id, values_from = value) %>%
    mutate(sig_label = case_when(
      P<0.05 & P>0.01 ~ "*", 
      P<0.01 & P>0.001 ~ "**", 
      P<0.001 ~ "***",
      TRUE ~ " "))}

data_spearman <- formatted_cors(targetdata)

data_spearman = as.data.frame(data_spearman)

colnames(data_spearman) = c("Y", "X", "Estimate", "N", "P_value", "sig_label")

data_spearman <- data_spearman %>% filter(Y == "GA" | Y == "BW" |  Y == "SIMD" | Y == "days_nnu")

data_spearman <- data_spearman %>% filter(X == "HD_Raw" | X == "FNF_Raw" |  X == "AA_Raw")

data_spearman <- data_spearman[-7,] #delete the correlation between BW and HD scale (run using Pearson because of data distribution, see below)

# pearson correlations

targetdata=Data_pret[,grep('BW|HD_', colnames(Data_pret))]
targetdata=targetdata[,-grep('BW_zscore', colnames(targetdata))]

cors <- function(df) { 
  # turn all three matrices (r, n, and P into a data frame)
  M <- Hmisc::rcorr(as.matrix(df), type="pearson")
  # return the three data frames in a list return(Mdf)
  Mdf <- map(M, ~data.frame(.x))
}

formatted_cors <- function(df){
  cors(df) %>%
    map(~rownames_to_column(.x, var="measure1")) %>%
    map(~pivot_longer(.x, -measure1, names_to="measure2")) %>% 
    bind_rows(.id = "id") %>%
    pivot_wider(names_from = id, values_from = value) %>%
    mutate(sig_label = case_when(
      P<0.05 & P>0.01 ~ "*", 
      P<0.01 & P>0.001 ~ "**", 
      P<0.001 ~ "***",
      TRUE ~ " "))}

data_pearson <- formatted_cors(targetdata)

data_pearson = as.data.frame(data_pearson)

colnames(data_pearson) = c("Y", "X", "Estimate", "N", "P_value", "sig_label")

data_pearson <- data_pearson %>% filter(Y == "BW")

data_pearson <- data_pearson %>% filter(X == "HD_Raw" )

# data wrangling and export

corr_cont_attachbeh <- rbind(data_spearman, data_pearson)

kable(corr_cont_attachbeh, col.names = gsub("[.]", " ", names(corr_cont_attachbeh)))

write.csv(corr_cont_attachbeh, here("results/covariates", "AttachBeh.cont.csv"), row.names = FALSE)

```

## Section III: Covariate analyses for secure vs insecure attachment classification

### Categorical and continuous variables

```{r}

# data wrangling

targetdata=Data_pret[,grep('AttachmentSecurity|BW|GA|SIMD|days_nnu|broncho', colnames(Data_pret))]
targetdata=targetdata[,-grep('BW_zscore|visit', colnames(targetdata))]

Pval_fun <- function(x) style_pvalue(x, digits = 3)
Stat_fun <- function(x) style_sigfig(x, digits=3)

# create table with comparisons

corr_attachclass <- targetdata %>%
  tbl_summary(
    by = AttachmentSecurity,
    statistic = list(c("GA", "SIMD", "days_nnu") ~ c("{median} ({min}, {max})"), c("BW") ~ c("{mean} ({sd})"), all_categorical() ~ "{n} ({p}%)"),
    digits = list(all_categorical()~2, all_continuous()~3)
  ) %>%
  add_p(
    test = list(c("GA", "SIMD", "days_nnu") ~ "wilcox.test", c("BW") ~ "t.test", all_categorical() ~ "chisq.test"),
    pvalue_fun = Pval_fun
  ) %>%
  separate_p_footnotes() %>%
  modify_header(list(statistic ~ "**t-statistic**", p.value ~ "**p-value**")) %>%
  modify_fmt_fun(list(statistic ~ Stat_fun, p.value ~ Pval_fun))

corr_attachclass

corr_attachclass <- as_tibble(corr_attachclass)

colnames(corr_attachclass) <- c("Characteristic", "Insecure attachment", "Secure attachment", "t-statistic", "p-value")

write.csv(corr_attachclass, here("results/covariates", "AttachClass.cat.cont.csv"), row.names=FALSE)

```


```{r pressure, echo=FALSE}
markdownToHTML("BM_Attachment_cov.Rmd",output ="BM_Attachment_cov.html")
```

