---
title: "BM_Attachment_comp"
author: "LJS"
date: "2024-07-01"
output: html_document
---

# Libraries 

```{r}

library("rlang") #toolbox
library("dplyr") #data manipulation
library("here") #toolbox
library("tidyverse") #data manipulation
library("tidytext") #data manipulation
library("knitr") #rmarkdown
library("htmltools") #rmarkdown
library("markdown") #rmarkdown
library("httpuv") #rmarkdown
library("vtable") #summary tables
library("gtsummary") #summary tables
library("stringr") #data manipulation
library("car") ##levene tests, qqplot
library("rstatix") ##kruskal_effsize, wilcox_effsize
library("nnet") ##multinom
library("ggplot2") ##plotting

```

# Breast milk exposure during neonatal care and infant attachment

Avoid scientific notation.

```{r}

options(scipen = 999)

```

## Section I: Prepare data

### Merge individual spreadsheets

```{r}

BM_variables <- read.csv(here("raw_data", "IDs_BM.csv"), header=T, dec=".",sep=",")
Attachment_variables <- read.csv(here("raw_data", "IDs_WT_rawtimes_filtered.csv"), header=T, dec=".",sep=",")
Covariate_variables <- read.csv(here("raw_data", "IDs_covariates.csv"), header=T, dec=".",sep=",")
TEA_feeding <- read.csv(here("raw_data", "IDs_TEA_feeding.csv"), header=T, dec=".",sep=",")

Data <- merge(BM_variables, Attachment_variables, by="ID")
Data <- merge(Covariate_variables, Data, by="ID")
Data <- merge(TEA_feeding, Data, by="ID")
rownames(Data) <- Data$ID
Data$ID <- NULL

```

### Breast milk

Calculate the % of exclusively mother breast milk, % of exclusively donor breast milk, % of both mother+donor breast milk and % of exclusively any (mother or donor) breast milk during neonatal care.

```{r}

Data <- Data %>% mutate(prop_days_excl_mother_breastmilk = (days_excl_mbm)/days_nnu)
Data <- Data %>% mutate(prop_days_excl_donor_breastmilk = (days_excl_debm)/days_nnu)
Data <- Data %>% mutate(prop_days_excl_motherdonor_breastmilk = (days_excl_mbm_debm)/days_nnu)
Data <- Data %>% mutate(prop_days_excl_any_breastmilk = (days_excl_mbm + days_excl_debm + days_excl_mbm_debm)/days_nnu)
Data <- Data %>% mutate(prop_days_formula = (days_excl_formula)/days_nnu)
Data <- Data %>% mutate(breastmilk_75_term = case_when((Prematurity==0)~2,
(Prematurity==1 & prop_days_excl_any_breastmilk < 0.75)~0,
(Prematurity==1 & prop_days_excl_any_breastmilk >= 0.75)~1))
Data=Data[complete.cases(Data$breastmilk_75_term),]

```

### Attachment

Categorise infants as secure or insecure based on attachment style. 
Remove participants with an unscorable attachment style (n = 7).

```{r}

Data$AttachmentSecurity = "Secure"

for(i in 1:nrow(Data)){
  if(Data$AttachmentStyle[i]=="Avoidant"){
    Data$AttachmentSecurity[i] <- "Insecure"}
  if(Data$AttachmentStyle[i]=="Resistant"){
    Data$AttachmentSecurity[i] <- "Insecure"}
  if(Data$AttachmentStyle[i]=="Disorganised"){
    Data$AttachmentSecurity[i] <- "Insecure"}
  if(Data$AttachmentStyle[i]=="NotFit"){
    Data$AttachmentSecurity[i] <- "NotFit"}}

Data = Data[Data$AttachmentSecurity=="Secure" | Data$AttachmentSecurity=="Insecure",]

```

### Transform variables to factors

```{r}

factor_columns <- c("breastmilk_75_term", "Sex", "Prematurity", "Exclusive_BF_4.5",
                    "Postnatal_depression_4.5", "Early_onset_sepsis", 
                    "Late_onset_sepsis", "Medical_or_surgical_nec", 
                    "bronchopulmonary_dysplasia", "retinopathy_or_prematurity")

for (col in factor_columns) {
  Data[[col]] <- as.factor(Data[[col]])
}

```

## Section II: Attachment measures per group 

```{r}

targetdata <- Data[,grep("breastmilk_75_term|HD_|AA_|FNF_|AttachmentSecurity", colnames(Data))]

tbl_summary(
  targetdata,
  by = breastmilk_75_term,
  label = NULL,
  list(c("HD_Raw", "FNF_Raw", "AA_Raw") ~ c("{median} ({min}, {max})"), all_categorical() ~ "{n} ({p}%)"),
  digits = NULL,
  type = NULL,
  value = NULL,
  missing = NULL,
  missing_text = NULL,
  sort = NULL,
  percent = NULL,
  include = everything()
)

```

## Section III: Analysis of breast milk exposure and attachment behaviours

### Attachment measures per group

```{r}

targetdata <- Data[,grep("breastmilk_75_term|HD_|AA_|FNF_|AttachmentSecurity", colnames(Data))]

tbl_summary(
  targetdata,
  by = breastmilk_75_term,
  label = NULL,
  list(c("HD_Raw", "FNF_Raw", "AA_Raw") ~ c("{median} ({min}, {max})"), all_categorical() ~ "{n} ({p}%)"),
  digits = NULL,
  type = NULL,
  value = NULL,
  missing = NULL,
  missing_text = NULL,
  sort = NULL,
  percent = NULL,
  include = everything()
)

```

### Make a dataframe for results

```{r}

results_table <- data.frame(matrix(ncol = 7, nrow = 4))
x <- c("measure", "statistic", "effsize_general", "pval_1_general", "pval_lowvshigh", "pval_1_lowvsterm", "pval_1_highvsterm")
colnames(results_table) <- x

results_table$measure <- c("HD_Raw", "FNF_Raw", "AA_Raw", "Secure/Insecure")

```

### Breast milk exposure ~ Happy-Distressed scale

```{r}

res_aov <- aov(HD_Raw ~ breastmilk_75_term, data = Data)
summary(res_aov)

```

#### Check model assumptions

```{r}

# Normality of residuals

shapiro.test(res_aov$residuals)
hist(res_aov$residuals)
qqPlot(res_aov$residuals,
  id = FALSE # id = FALSE to remove point identification
)

# Homogeneity of variance

leveneTest(res_aov)
plot(res_aov, which = 3)

```

The assumption of normal distribution of residuals is not met so Kruskal Wallis is used instead.

#### Kruskal Wallis

```{r}

kruskal_test <- kruskal.test(HD_Raw ~ breastmilk_75_term, data = Data)
kruskal_test

kruskal_effsize <- kruskal_effsize(
  Data,
  HD_Raw ~ breastmilk_75_term,
  ci = FALSE,
  conf.level = 0.95,
  ci.type = "perc",
  nboot = 1000
)

pairwise <- pairwise.wilcox.test(Data$HD_Raw, Data$breastmilk_75_term, p.adjust.method = "BH", pool.sd = FALSE)
pairwise

wilcox_effsize(HD_Raw~breastmilk_75_term, data=Data)

# Main p-values to be corrected for analyses of attachment behaviours and secure vs insecure attachment classification are labelled "pval_1_" .

results_table$statistic[which(results_table$measure=="HD_Raw")] <- kruskal_test$statistic
results_table$effsize_general[which(results_table$measure=="HD_Raw")] <- kruskal_effsize$effsize
results_table$pval_1_general[which(results_table$measure=="HD_Raw")] <- kruskal_test$p.value
results_table$pval_lowvshigh[which(results_table$measure=="HD_Raw")] <- pairwise[["p.value"]][1]
results_table$pval_lowvsterm[which(results_table$measure=="HD_Raw")] <- pairwise[["p.value"]][2]
results_table$pval_highvsterm[which(results_table$measure=="HD_Raw")] <- pairwise[["p.value"]][2,2]

```

### Breast milk exposure ~ Not fretful-Fretful scale

```{r}

res_aov <- aov(FNF_Raw ~ breastmilk_75_term, data = Data)
summary(res_aov)

```

#### Check model assumptions

```{r}

# Normality of residuals

shapiro.test(res_aov$residuals)
hist(res_aov$residuals)
qqPlot(res_aov$residuals,
  id = FALSE # id = FALSE to remove point identification
)

#Homogeneity of variance

leveneTest(res_aov)
plot(res_aov, which = 3)

```

The assumption of normal distribution of residuals is not met so Kruskal Wallis is used instead.

#### Kruskal Wallis

```{r}

kruskal_test <- kruskal.test(FNF_Raw ~ breastmilk_75_term, data = Data)
kruskal_test

pairwise <- pairwise.wilcox.test(Data$FNF_Raw, Data$breastmilk_75_term, p.adjust.method = "BH", pool.sd = FALSE)
pairwise

kruskal_effsize <-kruskal_effsize(
  Data,
  FNF_Raw ~ breastmilk_75_term,
  ci = FALSE,
  conf.level = 0.95,
  ci.type = "perc",
  nboot = 1000
)

wilcox_effsize(FNF_Raw~breastmilk_75_term, data=Data)

# Main p-values to be corrected for analyses of attachment behaviours and secure vs insecure attachment classification are labelled "pval_1_" .

results_table$statistic[which(results_table$measure=="FNF_Raw")] <- kruskal_test$statistic
results_table$effsize_general[which(results_table$measure=="FNF_Raw")] <- kruskal_effsize$effsize
results_table$pval_1_general[which(results_table$measure=="FNF_Raw")] <- kruskal_test$p.value
results_table$pval_lowvshigh[which(results_table$measure=="FNF_Raw")] <- pairwise[["p.value"]][1]
results_table$pval_lowvsterm[which(results_table$measure=="FNF_Raw")] <- pairwise[["p.value"]][2]
results_table$pval_highvsterm[which(results_table$measure=="FNF_Raw")] <- pairwise[["p.value"]][2,2]

```

### Breast milk exposure ~ Attentive-Avoidant scale

```{r}

res_aov <- aov(AA_Raw ~ breastmilk_75_term, data = Data)
summary(res_aov)

```

#### Check model assumptions

```{r}

#Normality of residuals

shapiro.test(res_aov$residuals)
hist(res_aov$residuals)
qqPlot(res_aov$residuals,
  id = FALSE # id = FALSE to remove point identification
)

#Homogeneity of variance

leveneTest(res_aov)
plot(res_aov, which = 3)

```

The assumption of normal distribution of residuals is not met so Kruskal Wallis is used instead.

#### Kruskal Wallis

```{r}

kruskal_test <- kruskal.test(AA_Raw ~ breastmilk_75_term, data = Data)
kruskal_test

pairwise <- pairwise.wilcox.test(Data$AA_Raw, Data$breastmilk_75_term, p.adjust.method = "BH", pool.sd = FALSE)
pairwise

kruskal_effsize <- kruskal_effsize(
  Data,
  AA_Raw ~ breastmilk_75_term,
  ci = FALSE,
  conf.level = 0.95,
  ci.type = "perc",
  nboot = 1000
)


wilcox_effsize(AA_Raw~breastmilk_75_term, data=Data)

# Main p-values to be corrected for analyses of attachment behaviours and secure vs insecure attachment classification are labelled "pval_1_" .

results_table$statistic[which(results_table$measure=="AA_Raw")] <- kruskal_test$statistic
results_table$effsize_general[which(results_table$measure=="AA_Raw")] <- kruskal_effsize$effsize
results_table$pval_1_general[which(results_table$measure=="AA_Raw")] <- kruskal_test$p.value
results_table$pval_lowvshigh[which(results_table$measure=="AA_Raw")] <- pairwise[["p.value"]][1]
results_table$pval_lowvsterm[which(results_table$measure=="AA_Raw")] <- pairwise[["p.value"]][2]
results_table$pval_highvsterm[which(results_table$measure=="AA_Raw")] <- pairwise[["p.value"]][2,2]

```

### Plotting attachment behaviours

```{r}

gHD <- ggplot(Data, aes(x=breastmilk_75_term, y=HD_Raw, group=as.factor(breastmilk_75_term), color=as.factor(breastmilk_75_term))) +
  geom_violin(lwd=0.8, width=0.5, show.legend = FALSE) +
  geom_boxplot(outlier.shape = NA, lwd=0.8, width=0.2, show.legend = FALSE) + 
  geom_jitter(width = 0.2, aes(fill=factor(breastmilk_75_term)), show.legend = FALSE) +
  scale_color_manual(values=c("#EB7C69", "#FDA638", "#459395")) +
  xlab('') + ylab("Happy - Distressed raw score") +
  scale_x_discrete(limits=c("0", "1", "2"), labels=c("Preterm low BM", "Preterm high BM", "Full-term")) +
  scale_y_continuous(limits=c(-0.05,1.10), breaks=seq(0,1, by = 0.25)) +
  theme_bw(base_size=13)
gHD
ggsave(here("results/figures", "HD.comp.pdf"), plot = gHD, dpi=300)

gFNF <- ggplot(Data, aes(x=breastmilk_75_term, y=FNF_Raw, group=as.factor(breastmilk_75_term), color=as.factor(breastmilk_75_term))) +
  geom_violin(lwd=0.8, width=0.5, show.legend = FALSE, trim=FALSE) +
  geom_boxplot(outlier.shape = NA, lwd=0.8, width=0.2, show.legend = FALSE) + 
  geom_jitter(width = 0.2, aes(fill=factor(breastmilk_75_term)), show.legend = FALSE) +
  scale_color_manual(values=c("#EB7C69", "#FDA638", "#459395")) +
  xlab('') + ylab("Not fretful - Fretful raw score") +
  scale_x_discrete(limits=c("0", "1", "2"), labels=c("Preterm low BM", "Preterm high BM", "Full-term")) +
  scale_y_continuous(limits=c(-0.05,1.10), breaks=seq(0,1, by = 0.25)) +
  theme_bw(base_size=13)
gFNF
ggsave(here("results/figures", "FNF.comp.pdf"), plot = gFNF, dpi=300)

gAA <- ggplot(Data, aes(x=breastmilk_75_term, y=AA_Raw, group=as.factor(breastmilk_75_term), color=as.factor(breastmilk_75_term))) +
  geom_violin(lwd=0.8, width=0.5, show.legend = FALSE) +
  geom_boxplot(outlier.shape = NA, lwd=0.8, width=0.2, show.legend = FALSE) + 
  geom_jitter(width = 0.2, aes(fill=factor(breastmilk_75_term)), show.legend = FALSE) +
  scale_color_manual(values=c("#EB7C69", "#FDA638", "#459395")) +
  xlab('') + ylab("Attentive - Avoidant raw score") +
  scale_x_discrete(limits=c("0", "1", "2"), labels=c("Preterm low BM", "Preterm high BM", "Full-term")) +
  scale_y_continuous(limits=c(-0.05,1.10), breaks=seq(0,1, by = 0.25)) +
  theme_bw(base_size=13)
gAA
ggsave(here("results/figures", "AA.comp.pdf"), plot = gAA, dpi=300)

```

## Section IV: Analysis of breast milk exposure and secure versus insecure attachment classification 

```{r}

Data$breastmilk_75_term2 <- relevel(factor(Data$breastmilk_75_term), ref="2")
Data$AttachmentSecurity2 <- relevel(factor(Data$AttachmentSecurity), ref="Insecure")

multinom_test <- multinom(AttachmentSecurity2 ~ breastmilk_75_term2, data = Data)
summary(multinom_test)
z <- summary(multinom_test)$coefficients/summary(multinom_test)$standard.errors
z
p <- (1 - pnorm(abs(z), 0, 1)) * 2
p
exp(coef(multinom_test))

# Main p-values to be corrected for analyses of attachment behaviours and secure vs insecure attachment classification are labelled "pval_1_" .

results_table$statistic[which(results_table$measure=="Secure/Insecure")] <- NA
results_table$effsize_general[which(results_table$measure=="Secure/Insecure")] <- NA
results_table$pval_1_general[which(results_table$measure=="Secure/Insecure")] <- NA
results_table$pval_lowvshigh[which(results_table$measure=="Secure/Insecure")] <- NA
results_table$pval_lowvsterm[which(results_table$measure=="Secure/Insecure")] <- NA
results_table$pval_highvsterm[which(results_table$measure=="Secure/Insecure")] <- NA
results_table$pval_1_lowvsterm[which(results_table$measure=="Secure/Insecure")] <- p[2]
results_table$pval_1_highvsterm[which(results_table$measure=="Secure/Insecure")] <- p[3]

```

## Section V: P-value adjustment and save results

```{r}

write.csv(results_table, here("results/comparisons", "Comp.3group.unadj.csv"), row.names = FALSE)

results_table_adjusted = results_table%>%mutate(across(contains("pval_1_"),~ p.adjust(.x, method = "BH", n = 5)))
write.csv(results_table_adjusted, here("results/comparisons", "Comp.3group.adj.csv"), row.names = FALSE)

```


```{r pressure, echo=FALSE}
markdownToHTML("BM_Attachment_comp.Rmd",output ="BM_Attachment_comp.html")
```

